```{r fig-symptoms, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'COVID-19 symptom recovery.'}

paper_figures$symptoms$plot

```

__Figure \@ref(fig:fig-symptoms). COVID-19 symptom recovery.__



```{r fig-lft, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'Functional lung recovery.'}

paper_figures$lft$plot

```

__Figure \@ref(fig:fig-lft). Functional lung recovery.__



```{r fig-ct, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Radiological lung recovery.'}

paper_figures$ct$plot

```

__Figure \@ref(fig:fig-ct). Radiological lung recovery.__



```{r fig-cardio, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Cardiological recovery.'}

paper_figures$cardio$plot

```

__Figure \@ref(fig:fig-cardio). Cardiological recovery.__



```{r fig-correlations, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Persistent symptoms and cardiopulmonary abnormalities and mobility, health self-perception, fatigue and stress scoring.'}

paper_figures$correlations$plot

```

__Figure \@ref(fig:fig-correlations). Persistent symptoms and cardiopulmonary abnormalities and mobility, health self-perception, fatigue and stress scoring.__



```{r fig-clusters, fig.width = 7.086614166, fig.height = 8.661417314, fig.cap = 'Clusters of clinical and psychosocial COVID-19 recovery.'}

paper_figures$clusters$plot

```

__Figure \@ref(fig:fig-clusters). Clusters of clinical and psychosocial COVID-19 recovery.__



```{r fig-clust-psych, fig.width = 7.086614166, fig.height = 8.661417314, fig.cap = 'Quality of life, fatigue and mental health rating in the COVID-19 recovery clusters.'}

paper_figures$clust_psych$plot

```

__Figure \@ref(fig:fig-clust-psych). Quality of life, fatigue and mental health rating in the COVID-19 recovery clusters.__

