---
title: "COVID-19 and its continuing burden after 12 months: a longitudinal observational prospective multicenter trial"
subtitle: "Figures and Tables"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx
    
bibliography: cov_biblio.bib
csl: erj.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

  library(bookdown)

  knitr::opts_chunk$set(echo = FALSE, 
                        warning = FALSE, 
                        message = FALSE, 
                        dev = "png", 
                        dpi = 600)
  
  flextable::set_flextable_defaults(font.family = 'Cambria', 
                                    font.size = 10)


```

\newpage

# Tables

```{r tab-cohort, tab.cap = 'Baseline characteristics, post-acute steroid therapy and rehabilitation status of the study cohort and the COVID-19 severity groups.'}

paper_tables$cohort_chara %>% 
  map_dfc(stri_replace, fixed = 'no:', replacement = 'none:') %>% 
  flextable %>% 
  width(1, width = 3, unit = 'cm') %>% 
  width(2:5, width = 2.5, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 1.8, unit = 'cm') %>% 
  theme_vanilla %>% 
  footnote(i = 1, j = 6:7, value = as_paragraph('Comparison between the COVID-19 severity strata. Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic; numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>% 
  footnote(i = 4, j = 1, value = as_paragraph('Body mass index (BMI); overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2'), ref_symbols = 'b') %>% 
  footnote(i = 16, j = 1, value = as_paragraph('Steroid therapy in cases of non-resolving pneumonia beginning from week four post diagnosis at the discretion of the physician.'), ref_symbols = 'c')

```

\newpage

```{r tab-enpoints, tab.cap = 'Key outcome measures at the one-year follow-up in the study cohort and the COVID-19 severity groups.'}

paper_tables$endpoints %>% 
  flextable %>% 
  width(1, width = 3, unit = 'cm') %>% 
  width(2:5, width = 2.5, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 1.8, unit = 'cm') %>% 
  theme_vanilla %>% 
  footnote(i = 1, j = 6:7, value = as_paragraph('COVID-19 severity groups compared with \u03C7\u00B2 test with Cramer V effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>% 
  footnote(i = 2, j = 1, value = as_paragraph('LFT abnormality: abnormality in lung function testing, <80% predicted value (FEV1: forced expiratory volume in 1 second; FVC: forced vital capacity; DLCO: diffusion lung capacity for carbon monoxide; TLC: total lung capacity) or <70% predicted value cutoffs (FEV1:FVC: FEV1 to FVC ratio.)'), ref_symbols = 'b') %>% 
  footnote(i = 3, j = 1, value = as_paragraph('CT abnormality: any abnormality in chest computed tomography, CT severity score \u2265 1.'), ref_symbols = 'c')

```

\newpage

```{r tab-psych-reco, tab.cap = 'Physical performance, fatigue, self-perceived general health, quality of life and mental health readouts at the one-year follow-up in the study cohort and the COVID-19 severity groups.'}

paper_tables$psych_reco %>% 
  flextable %>% 
  width(1, width = 3, unit = 'cm') %>% 
  width(2:5, width = 2.5, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 1.8, unit = 'cm') %>% 
  theme_vanilla %>% 
  footnote(i = 1, j = 6:7, value = as_paragraph('Comparison between the COVID-19 severity strata. Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic; numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>% 
  footnote(i = 1, j = 1, value = as_paragraph('6MWD: six-minute walking distance, meters.'), ref_symbols = 'b', part = 'body') %>% 
  footnote(i = 2, j = 1, value = as_paragraph('6MWD vs ref.: difference between the reference and observed 6MWD value, meters'), ref_symbols = 'c', part = 'body') %>% 
  footnote(i = 3, j = 1, value = as_paragraph('6MWD < ref.: 6MWD below the reference value.'), ref_symbols = 'd', part = 'body') %>% 
  footnote(i = 4:5, j = 1, value = as_paragraph('CFS: 11-item Chalder fatigue score, incr.: increased.'), ref_symbols = 'e', part = 'body') %>% 
  footnote(i = 6:7, j = 1, value = as_paragraph('EQ5D5L: European quality of life 5 dimensions, 5 levels, VAS: visual analogue scale, imp.: impaired.'), ref_symbols = 'f', part = 'body') %>% 
  footnote(i = 8:17, j = 1, value = as_paragraph('EQ5D5L: European quality of life 5 dimensions, 5 levels, imp.: impaired'), ref_symbols = 'g', part = 'body') %>% 
  footnote(i = 18:19, j = 1, value = as_paragraph('PSS: 4-item perceived stress scale.'), ref_symbols = 'h', part = 'body') %>% 
  footnote(i = 20, j = 1, value = as_paragraph('SSD-12: 12-item somatic syndrome disorder – B criteria scale.'), ref_symbols = 'i', part = 'body') %>% 
  footnote(i = 21, j = 1, value = as_paragraph('BRCS: brief resilient coping score.'), ref_symbols = 'j', part = 'body') %>% 
  footnote(i = 22, j = 1, value = as_paragraph('Resilient coping class: low: 4 - 13 points, medium: 14 - 16 points, high: 17 - 21 points of the BRCS scale.'), ref_symbols = 'k', part = 'body')

```


\newpage

# Figures {#sec:figures}

```{r fig-consort, fig.width = mm_inch(paper_figures$consort$w), fig.height = mm_inch(paper_figures$consort$h), dpi = 600, fig.cap = 'Flow diagram of the study analysis inclusion.'}

paper_figures$consort$plot

```

__Figure \@ref(fig:fig-consort). Flow diagram of the study analysis inclusion.__

\newpage

```{r fig-symptoms, fig.width = mm_inch(paper_figures$symptoms$w), fig.height = mm_inch(paper_figures$symptoms$h), fig.cap = 'COVID-19 symptom recovery.'}

paper_figures$symptoms$plot

```

__Figure \@ref(fig:fig-symptoms). COVID-19 symptom recovery.__

_Presence of COVID-19 symptoms (reduced performance: Eastern Cooperative Oncology group score [ECOG] $\geq$ 1, fatigue: bimodal Chalder fatigue score [CFS] $\geq$ 4, dyspnea: modified Medical Research Council score [mMRC] $\geq$ 1, self reported: sleep problems, night sweating, cough, hair loss, hyposmia/anosmia, dermatological and gastrointestinal symptoms) was analyzed in the entire study collective and in ambulatory, moderate and severe COVID-19 survivors._

_(A) Percentages of individuals with particular symptoms at the one-year follow-up. Differences between the COVID-19 severity strata were investigated by $\chi^2$ test corrected for multiple testing with Benjamini-Hochberg method. Point size and color code for the percentage. P values are displayed in the Y axis. Numbers of complete observations are indicated in the X axis._

_(B) Percentages of individuals with any symptoms during acute COVID-19 and at the 2-, 3-, 6-month and one-year follow-up. Participants with the complete longitudinal data set were included in the analysis. The symptom kinetic was analyzed by second-order mixed-effect logistic modeling and likelihood ratio test (full vs null model). P values were corrected for multiple testing with the Benjamini-Hochberg method. Likelihood ratio ($\lambda$), p values and numbers of participants with the complete longitudinal data set are presented in the plot captions._

\newpage

```{r fig-lft, fig.width = mm_inch(paper_figures$lft$w), fig.height = mm_inch(paper_figures$lft$h), fig.cap = 'Functional lung recovery.'}

paper_figures$lft$plot

```

__Figure \@ref(fig:fig-lft). Functional lung recovery.__

_Lung function testing (LFT) was analyzed in the entire study collective and in ambulatory, moderate and severe COVID-19 survivors._

_(A) Percentages of individuals with particular LFT abnormalities at the one-year follow-up. Differences between the COVID-19 severity strata were investigated by $\chi^2$ test corrected for multiple testing with Benjamini-Hochberg method. Point size and color code for the percentage. P values are displayed in the Y axis. Numbers of complete observations are indicated in the X axis. FEV1: forced expiratory volume in 1 second; FVC: forced vital capacity; DLCO: diffusion lung capacity for carbon monoxide; RV: residual volume, FEV1:FVC: FEV1 to FVC ratio, TLC: total lung capacity._

_(B) Percentages of individuals with any LFT abnormality (FEV1, FVC, DLCO or TLC: <80% predicted value or FEV1:FVC <70% predicted value) at the 2-, 3-, 6-month and one-year follow-up. Participants with the complete longitudinal data set were included in the analysis. The LFT finding kinetic was analyzed by second-order mixed-effect logistic modeling and likelihood ratio test (full vs null model). P values were corrected for multiple testing with the Benjamini-Hochberg method. Likelihood ratio ($\lambda$), p values and numbers of participants with the complete longitudinal data set are presented in the plot captions._

\newpage

```{r fig-ct, fig.width = mm_inch(paper_figures$ct$w), fig.height = mm_inch(paper_figures$ct$h), fig.cap = 'Radiological lung recovery.'}

paper_figures$ct$plot

```

__Figure \@ref(fig:fig-ct). Radiological lung recovery.__

_Chest computed tomography (CT) was analyzed in the entire study collective and in ambulatory, moderate and severe COVID-19 survivors._

_(A) Percentages of individuals with any chest CT abnormality (CT severity score $\geq$ 1) and abnormalities scored >5 CT severity score points at the one-year follow-up. Differences between the COVID-19 severity strata were investigated by $\chi^2$ test corrected for multiple testing with Benjamini-Hochberg method. Point size and color code for the percentage. P values are displayed in the Y axis. Numbers of complete observations are indicated in the X axis._

_(B) Percentages of individuals with any chest CT abnormality at the 2-, 3-, 6-month and one-year follow-up. Participants with the complete longitudinal data set were included in the analysis. The CT finding kinetic was analyzed by second-order mixed-effect logistic modeling and likelihood ratio test (full vs null model). P values were corrected for multiple testing with the Benjamini-Hochberg method. Likelihood ratio ($\lambda$), p values and numbers of participants with the complete longitudinal data set are presented in the plot captions._

\newpage

```{r fig-cardio, fig.width = mm_inch(paper_figures$cardio$w), fig.height = mm_inch(paper_figures$cardio$h), fig.cap = 'Cardiological recovery.'}

paper_figures$cardio$plot

```

__Figure \@ref(fig:fig-cardio). Cardiological recovery.__

_Transthoracic echocardiography (TTE) was performed in the entire study collective and in ambulatory, moderate and severe COVID-19 survivors._

_(A) Percentages of individuals diagnosed diastolic dysfunction of any severity and reduced left ventricular ejection fraction (LVEF) at the one-year follow-up. Differences between the COVID-19 severity strata were investigated by $\chi^2$ test corrected for multiple testing with Benjamini-Hochberg method. Point size and color code for the percentage. P values are displayed in the Y axis. Numbers of complete observations are indicated in the X axis._

_(B) Percentages of individuals diagnosed diastolic dysfunction at the 2-, 3-, 6-month and one-year follow-up. Participants with the complete longitudinal data set were included in the analysis. The diastolic dysfunction kinetic was analyzed by second-order mixed-effect logistic modeling and likelihood ratio test (full vs null model). P values were corrected for multiple testing with the Benjamini-Hochberg method. Likelihood ratio ($\lambda$), p values and numbers of participants with the complete longitudinal data set are presented in the plot captions._

\newpage

```{r fig-correlations, fig.width = mm_inch(paper_figures$correlations$w), fig.height = mm_inch(paper_figures$correlations$h), fig.cap = 'Correlation of symptoms, physical performance, cardiopulmonary findings, mental health and quality of life at the one-year follow-up.'}

paper_figures$correlations$plot

```

__Figure \@ref(fig:fig-correlations). Correlation of symptoms, physical performance, cardiopulmonary findings, mental health and quality of life at the one-year follow-up.__

_Association of COVID-19 symptoms (number of symptoms: # symptoms, dyspnea score: modified Medical Research Council score [mMRC], reduced performance: Eastern Cooperative Oncology Group score [ECOG] $\geq$ 1, fatigue: likert Chalder fatigue score [CFS], self-reported: sleep problems, cough, night sweating and hyposmia/anosmia), mobility (6MWD < ref.: six minute walking distance, difference versus the reference value), lung function (LFT: any lung function testing abnormality, DLCO: diffusion lung capacity for carbon monoxide), chest computed tomography (CT) severity score, diastolic dysfunction, self-perceived general health (EQ5D5L VAS: European quality of life 5 dimensions, 5 levels, visual analogue scale), quality of life and mental health scoring (EQ5D5L: European quality of life 5 dimensions, 5 levels) and stress (PSS: 4-item perceived stress scale) at the one-year follow-up. Pairwise correlations were investigated by Kendall's $\tau$ B test. P values were corrected for multiple testing with the Benjamini-Hochberg method. $\tau$ coefficients for significant correlations are presented. Point size and color code for the $\tau$ value. The number of complete observations is indicated in the plot caption._

\newpage

```{r fig-clusters, fig.width = mm_inch(paper_figures$clusters$w), fig.height = mm_inch(paper_figures$clusters$h), fig.cap = 'COVID-19 recovery clusters.'}

paper_figures$clusters$plot

```

__Figure \@ref(fig:fig-clusters). COVID-19 recovery clusters.__

_Clustering of the study participants in respect to symptoms (any symptom present, dyspnea: modified Medical Research Council score [mMRC] $\geq$ 1, reduced performance: Eastern Cooperative Oncology Group score [ECOG] $\geq$ 1, fatigue: bimodal Chalder fatigue score [CFS] $\geq$ 4, self-reported: sleep problems, cough, night sweating and hyposmia/anosmia), mobility (6MWD < ref.: six minute walking distance, difference versus the reference value), cardiopulmonary abnormalities (any chest computed tomography [CT] abnormality: CT severity score $\geq$ 1, any lung function testing [LFT] abnormality, diastolic dysfunction), significant stress (4-item perceived stress scale [PSS] >5), impaired self-perceived general health (European quality of life 5 dimensions, 5 levels, visual analogue scale [EQ5D5L VAS] <73, imp.: impaired) as well as features of quality of life and mental health (European quality of life 5 dimensions, 5 levels [EQ5D5L], cutoff: score >1) at the one-year follow-up. Clustering analysis was done with the PAM algorithm (PAM: partitioning around medoids, simple matching distance). Differences in frequency of the cardiopulmonary (A), symptom and mobility (B) as well as self-perceived general health, quality of life and mental health (C) clustering variables between the recovery clusters were analyzed by $\chi^2$ test. P values were corrected for multiple testing with the Benjamini-Hochberg method. Lines represent the estimated percentages of the feature in the cluster, tinted regions represent 2$\times$SEM intervals. P values are indicated in the Y axis. Numbers of participants assigned to the clusters are displayed in A._

\newpage

```{r fig-clust-psych, fig.width = mm_inch(paper_figures$clust_psych$w), fig.height = mm_inch(paper_figures$clust_psych$h), fig.cap = 'Quality of life, fatigue and mental health rating in the COVID-19 recovery clusters.'}

paper_figures$clust_psych$plot

```

__Figure \@ref(fig:fig-clust-psych). Quality of life, fatigue and mental health rating in the COVID-19 recovery clusters.__

_Quality of life, general health and rating of fatigue, stress, somatic symptom disorder and resilience were assessed at the one-year follow-up were compared between the COVID-19 recovery clusters. Statistical significance was determined by Kruskal-Wallis test with $\eta^2$ effect size statistic. P values were corrected for multiple testing with Benjamini-Hochberg method. Effect size statistic and p values are presented in plot captions. Numbers of participants assigned to the clusters are presented under the plots._

_EQ5D5L: European quality of life 5 dimensions, 5 levels; VAS: visual analogue scale; CFS: Chalder fatigue score; SSD-12: somatic syndrome disorder - B criteria scale; PSS: 4-item perceived stress scale; BRCS: brief resilient coping scale._

\newpage