---
title: "COVID-19 and its continuing burden after 12 months: a longitudinal observational prospective multicenter trial"
subtitle: "Supplementary Material"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx
    
bibliography: cov_biblio.bib
csl: erj.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

  library(bookdown)
  
  knitr::opts_chunk$set(echo = FALSE, 
                        warning = FALSE, 
                        message = FALSE, 
                        dev = "png", 
                        dpi = 600)
  
  flextable::set_flextable_defaults(font.size = 10, 
                                    font.family = 'Cambria')


```

\newpage

\beginsupplement

# Supplementary Methods

## Software

The study variables were transformed, analyzed and visualized with R version 4.2.0. 
General data import and transformation tasks were accomplished with the _tidyverse_ [@Wickham2019] and _rlang_ [@Henry2022] packages. 
Analysis results were visualized with the _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019], _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022] and _bookdown_ [@Xie2016] packages and the development package _figur_ (https://github.com/PiotrTymoszuk/figur). 
Tables were generated with the package _flextable_ [@Gohel2022].

## Procedures, variables and variable stratification

Numeric variables are presented as medians with interquartile ranges (IQR), categorical variables are presented as percentages of complete answers.

Participant age during acute COVID-19 was stratified for modeling tasks with the 50 and 65 year cutoffs. 
Participants were stratified according to the severity of acute COVID-19 as ambulatory (outpatients), moderate (hospitalized at COVID-19 ward, no mechanical ventilation) and severe (mechanical ventilation and/or intensive care unit stay).

Investigated COVID-19 symptoms encompassed 
self-reported complaints (fever, night sweating, cough, smell disorders, sleep disorders, hair loss, dermatological and gastrointestinal symptoms, surveyed by single yes/no questions), 
dyspnoea (Modified Medical British Research Council [mMRC], dyspnea: mMRC $\geq$ 1), 
reduced physical performance (Eastern Cooperative Oncology Group [ECOG], reduced performance: ECOG ≥1) [@Oken1982] 
and fatigue (bimodal and likert 11-item Chalder fatigue score [CFS]; fatigue: bimodal CFS $\geq$ 4) [@Morriss1998].

Lung function testing (LFT) parameters were stratified by 80% predicted value (FEV1: forced expiratory volume in 1 second; FVC: forced vital capacity; DLCO: diffusion lung capacity for carbon monoxide; TLC: total lung capacity) or 70% predicted value cutoffs (FEV1:FVC: FEV1 to FVC ratio). Abnormal LFT was diagnosed when at least one of FEV1, FVC, DLCO, RV, TLC or FEV1:FVC was reduced. 

Chest computed tomography was assessed according to the Fleischner society glossary terms [@Hansell2008] and rated with the CT severity score summed over all lung lobes [@Sonnweber2022; @Sonnweber2020; @Luger2022].
The CT score for each lobe was defined as follows: 
0: none; 
1: minimal (subtle ground-glass opacities [GGO]); 
2: mild (several GGOs, subtle reticulation); 
3: moderate (multiple GGOs, reticulation, small consolidation); 
4: severe (extensive GGOs, consolidation, reticulation with distortion); 
and 5: massive (massive findings, parenchymal destruction). 
The total CT severity score was the sum overa all lung lobes. 
Any chest CT abnormalities were defined as CT severity score $\geq$ 1, moderate-to-severe CT abnormalities were defined as CT severity score > 5.

TTE was performed according to the ESC and EACVI recommendations [@Lancellotti2015].
Diastolic dysfunction was determined by assessing left ventricular filling by trans-mitral PW Doppler and measuring the peak of early filling (E) and peak of late atrial filling (A). 
For the assessment of pseudo-normal E/A ratio, peak early diastolic velocity (e’) was measured at the lateral mitral annulus with tissue Doppler to calculate E/e’. 
Grading of diastolic dysfunction was performed as follows: normal diastolic function: E/A ratio = 1 - 2; diastolic dysfunction I°: E/A ratio < 0.8; diastolic dysfunction II°: E/A = 0.8 - 1.5. and diastolic dysfunction III°: E/A > 2. 

Reference values of the six-minute walking distance (6MWD) [@Crapo2012] were calculated with the participants sex, age, weight and height as described previously [@Enrichi2012]. 
Low 6MWD was defined as a value below the patient's specific reference value.

Sub-scores of the EQ-5D-5L questionnaire (European quality of life 5 dimensions, 5 levels) [@Herdman2011] addressing impairment of usual activities, mobility and self-care as well as pain/discomfort, depression/anxiety were stratified with the cutoff of 1 (1: no impairment or absence, > 1: impairment or presence). 
Visual analogue scale (VAS) of the EQ-5D-5L tool describing the self-perceived general health was binarized with the cutoff of 73.2 as published before for the general elderly German population [@Marten2021]. 
Elevated stress levels were defined by the median split of the 4-item PSS score (Perceived Stress Scale, > 5: elevated stress) [@Cohen1983]. 
Resilient coping classes: low (4 - 13 points), medium (14 - 16) and high (17 - 21 points) were defined with the Brief Resilient Coping Scale (BRCS) [@Sinclair2004].

Laboratory parameters were determined by the ISO-certified central laboratory of the University Hospital of Innsbruck (Zentralinstitut für medizinische u. chemische Labordiagnostik, ZIMCL).
C-reative protein (CRP), interleukin 6 (IL6), procalcitonin (PCT), N-terminal pro-brain natriuretic peptide (NT-proBNP) and serum ferritin (FT) were determined with a Roche Cobas 8000 analyser (Basel, Switzerland). 
D-dimer was measured with a Siemens BCS-XP instrument using the Siemens D-Dimer Innovance reagent (Erlangen, Germany). 
Anti-S1/S2 protein SARS-CoV-2 immunoglobulin gamma (IgG) was quantified with LIAISON chemoluminescence assay (DiaSorin, Italy) and expressed as arbitrary units.
Anemia was defined as hemoglobin < 140 g/dL (male) or < 120 g/dL (female). 
Elevated FT was defined as $\geq$ 300 µg/L or $\geq$ 150 µg/L for men and women, respectively. 
Reduced transferrin saturation (TSAT) was defined as < 20% or < 15% for men and women, respectively. 
Elevated NT-proBNP was defined as > 150 pg/mL. 
Elevated D-dimer, CRP, PCT and IL6 were defined with the 500 µg/L, 0.5 mg/L, 0.15 µg/L and 7 pg/mL cutoffs, respectively. 
Anti-S1/S2 SARS-CoV-2 immunoglubulin gamma (anti-S1/S2 IgG) was stratified by quartiles (cutoffs: `r quantile(cov_data$data_tbl$SarsCov2_IgG, c(0.25, 0.5, 0.75), na.rm = TRUE) %>% paste(collapse = ', ')` arbitrary units) [@Sonnweber2022].

The full list of the study variables and their stratification scheme is presented in __Supplementary Table\@ref(tab:tab-study-vars)__.

## Hypothesis testing and correlation

To compare differences in distribution of categorical features, $\chi^2$ test with Cramer V effect size statistic was applied. 
Since multiple numeric variables were strongly non-normally distributed as identified by Shapiro-Wilk test, Mann-Whitney U test with r effect size statistic or Kruskal-Wallis test with $\eta^2$ effect size statistic was applied to assess differences between two or more groups, respectively. 
Associations of the readouts of clinical, cardiopulmonary, mobility and quality of life deficits at the one-year follow-up visit were determined by pairwise Kendall's $\tau$ B correlation. 
In the correlation analysis, binary variables were recoded as 1 and 2 for absent and present, respectively. 
P values were corrected for multiple comparisons with Benjamini-Hochberg method [@Benjamini1995]. 
R packages _rstatix_ [@Kassambara2021], _rcompanion_ [@Mangiafico2022] and _Hmisc_ [@Harrell2022] and the in-house developed package _ExDA_ (https://github.com/PiotrTymoszuk/ExDA) were used for explorative data analysis, statistical hypothesis testing and correlation analysis.

## Modeling of risk and parameter value in time

To model recovery kinetics for categorical variables, second-order mixed-effect logistic (categorical features) modeling was applied (packages: _lme4_ and _lmerTest_) [@Bates2015; @Kuznetsova2017; @Box1978]. 
Each model followed the general formula:

$$Response \sim time + time^2 + (1|individual)$$

where $(1|individual)$ indicates the random effect of the individual and $time$ and $time^2$ indicate the first- and second-order time effect terms. 
The first-order term estimate was interpreted as a measure of the recovery speed and the second-order term estimate was used to assess the plateau/rebound effect. 
Significance of the accuracy gain of the full second-order model compared with the nested null model was determined by likelihood ratio test (LRT) versus the nested first-order and null models, respectively. 
The likelihood ratio $\lambda$ statistic was used as an effect size measure. 
Results of the kinetic modeling were adjusted for multiple comparisons with Benjamini-Hochberg method [@Benjamini1995].

## Clustering analysis

To identify patterns of COVID-19 recovery defined by `r length(globals$clust_variables)` binary symptom, cardiopulmonary and psychosocial variables recorded at the one-year follow-up visit (__Supplementary Tables S\@ref(tab:tab-study-vars)__) the subset of the study participants with the complete variable record was clustered using the PAM (partition around medoids) algorithm and the simple matching distance statistic (packages _cluster_ and _nomclust_) [@Schubert2019; @Boriah2008; @Sulc2021]. 
The data pre-processing included conversion of binary features to the numeric format (absent: 1, present: 1). 
The choice of the clustering procedure was motivated by the analysis of the clustering variance (ratio of the total between-cluster to total sum of squares) and clustering structure stability in 10-fold cross-validation (metric: rate of correct cluster assignment, cluster assignment predicted by k = 5 nearest neighbors label propagation algorithm, packages _dbscan_ and the development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Lange2004; @Leng2014] for several clustering algorithms as presented in __Supplementary Figure S\@ref(fig:fig-clust-qc)A__.
The optimal number of clusters was determined by the bend of the total within-cluster sum of squares curve (package _factoextra_) [@Kassambara2020] as presented in __Supplementary Figure S\@ref(fig:fig-clust-qc)B__.
Permutation importance of the clustering factors was assessed by comparing the clustering variance (ratio of the total between-cluster to total sum of squares) of the original clustering structure with the clustering variance of the clustering object with the given variable reshuffled randomly. 
The difference in clustering variances ($\Delta$ clustering variance) served as the importance statistic. 
Determination of the clustering factor importance was done for n = 100 permutations for each variable using the development package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools) as presented in __Supplementary Figure \@ref(fig:fig-clust-qc)C__.

## Multi-parameter modeling

The risk of persistent COVID-19 symptoms, LFT findings, CT abnormalities, and diastolic dysfunction at the one-year follow-up was modeled with multi-parameter logistic LASSO (least absolute shrinkage and selection operator) regression [@Tibshirani1996] and `r length(multi_mod$variables)` independent variables recorded during acute COVID-19 and at the 60-day follow-up (__Supplementary Table S\@ref(tab:tab-study-vars)__). 
The LASSO modeling was accomplished with the packages _glmnet_ [@Friedman2010] and _caret_ [@Kuhn2008].
The value of the $\lambda$ shrinkage parameter were determined by 10-fold cross-validation (CV) repeated 100 times; the optimal value for the model with the lowest deviance was chosen [@Friedman2010]. 
The model performance at predicting the one-year event in the training data set and in 10-fold CV was investigated by receiver-operating characteristic (ROC) [@Kuhn2008], Cohen's $\kappa$ [@Fleiss1969] and R-squared statistics (package _caret_) [@Kuhn2008]. 
Model assumptions were checked by a visual control of model residuals plots (residuals versus fitted and quantile-quantile plots, development package _caretExtra_ [https://github.com/PiotrTymoszuk/caretExtra]). 
ROC plots were generated with the _plotroc_ package [@Sachs2017] and the development package _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra).

## Source code availability

The raw study data will be made available upon request. The R analysis pipeline is available at https://github.com/PiotrTymoszuk/CovILD-Plus.

\newpage

# Supplementary Tables {#sec:tables}

```{r tab-study-vars, tab.cap = 'Study variables. The table is available as a supplementary Excel file.'}

suppl_tables$study_vars[1, 1] %>% 
  mutate(`R variable` = ' ') %>% 
  set_names(' ') %>% 
  flextable

```

\newpage

```{r tab-clust-base, tab.cap = 'Extended baseline characteristics, post-acute steroid therapy and rehabilitation status of the COVID-19 recovery clusters.'}

suppl_tables$clust_chara %>% 
  map_dfc(stri_replace, fixed = 'no', replacement = 'none') %>% 
  flextable %>% 
  width(1, width = 2.8, unit = 'cm') %>% 
  width(2:4, width = 4.3, unit = 'cm') %>% 
  width(5, width = 2.7, unit = 'cm') %>% 
  width(6, width = 2.3, unit = 'cm') %>%
  footnote(i = 1, j = 5:6, value = as_paragraph('\u03C7\u00B2 test with Cramer V effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>% 
  footnote(i = 11, j = 1, value = as_paragraph('Steroid therapy in cases of non-resolving pneumonia beginning from week four post diagnosis at the discretion of the physician.'), ref_symbols = 'b', part = 'body') %>% 
  theme_vanilla

```

\newpage

```{r tab-clust-fup, tab.cap = 'Extended characteristic of the COVID-19 recovery clusters at the one-year follow-up.'}

flextable::flextable(suppl_tables$clust_fup) %>% 
  width(1, width = 2.6, unit = 'cm') %>% 
  width(2:4, width = 4.3, unit = 'cm') %>% 
  width(5, width = 2.7, unit = 'cm') %>% 
  width(6, width = 2.3, unit = 'cm') %>%
  footnote(i = 1, j = 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic; numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>% 
  footnote(i = 5, j = 1, part = 'body', value = as_paragraph('CT severity score: chest computed tomography severity score'), ref_symbols = 'b') %>% 
  footnote(i = 6, j = 1, part = 'body', value = as_paragraph('6MWD, m: six-minute walking distance, meters'), ref_symbols = 'c') %>% 
  footnote(i = 7:8, j = 1, part = 'body', value = as_paragraph('EQ5D5L: European quality of life 5 dimensions, 5 levels'), ref_symbols = 'd') %>% 
  theme_vanilla

```


\newpage

# Supplementary Figures {#sec:figures}

```{r fig-symptoms1, fig.width = mm_inch(suppl_figures$symptoms1$w), fig.height = mm_inch(suppl_figures$symptoms1$h), fig.cap = 'Recovery of fatigue, sleep problems and dyspnea.'}

suppl_figures$symptoms1$plot

```

__Supplementary Figure S\@ref(fig:fig-symptoms1). Recovery of fatigue, sleep problems and dyspnea.__ 

_Frequencies of reduced physical performance (Eastern Cooperative Oncology score [ECOG] $\geq$ 1) (A), self-reported sleep problems (B) and dyspnea (modified Medical Research Council score [mMRC] $\geq$ 1) (C) during acute COVID-19 and at the 2-, 3-, 6-month and one-year follow-up were investigated in the entire study collective and in ambulatory, moderate and severe COVID-19 survivors. Participants with the complete longitudinal data set were included in the analysis. The symptom kinetic was analyzed by second-order mixed-effect logistic modeling and likelihood ratio test (full vs null model). P values were corrected for multiple testing with the Benjamini-Hochberg method. Likelihood ratio ($\lambda$), p values and numbers of participants with the complete longitudinal data set are presented in the plot captions._

\newpage

```{r fig-overlap1, fig.width = mm_inch(suppl_figures$phys_mental_recovery1$w), fig.height = mm_inch(suppl_figures$phys_mental_recovery1$h), fig.cap = 'Impact of persistent symptoms and LFT abnormality on physical and mental recovery from COVID-19.'}

suppl_figures$phys_mental_recovery1$plot

```

__Supplementary Figure S\@ref(fig:fig-overlap1). Impact of persistent symptoms and LFT abnormality on physical and mental recovery from COVID-19.__ 

_Six-minute walking distance (6MWD vs ref.: difference versus the reference value), fatigue rating (CFS: Chalder fatigue score), self-perceived general health (EQ5D5L VAS: European quality of life 5 dimensions, 5 levels, visual analogue scale), self-reported mobility, self-care and activity impairment scores (EQ5D5L: European quality of life 5 dimensions, 5 levels, imp.: impairment), pain/discomfort and depression/anxiety ratings (EQ5D5L), stress (PSS: 4-item perceived stress scale), resilience (BRCS: brief resilient coping scale), and somatic symptom disorder scoring (SSD-12: somatic symptom disorder - B criteria scale) were compared between participants with or without persistent symptoms (A) and with or without lung function testing (LFT) abnormalities (B) at the one-year follow-up. Statistical significance was assessed by Mann-Whitney U test corrected for multiple testing with Benjamini-Hochberg method. Normalized (Z score) mean variable values with 2$\times$SEM are presented in radial plots. Numbers of complete observations are indicated under the plots._

\newpage

```{r fig-overlap2, fig.width = mm_inch(suppl_figures$phys_mental_recovery2$w), fig.height = mm_inch(suppl_figures$phys_mental_recovery2$h), fig.cap = 'Impact of lung CT and cardiological abnormality on physical and mental recovery from COVID-19.'}

suppl_figures$phys_mental_recovery2$plot

```

__Supplementary Figure S\@ref(fig:fig-overlap2). Impact of lung CT and cardiological abnormality on physical and mental recovery from COVID-19.__ 

_Six-minute walking distance (6MWD vs ref.: difference versus the reference value), fatigue rating (CFS: Chalder fatigue score), self-perceived general health (EQ5D5L VAS: European quality of life 5 dimensions, 5 levels, visual analogue scale), self-reported mobility, self-care and activity impairment scores (EQ5D5L: European quality of life 5 dimensions, 5 levels, imp.: impairment), pain/discomfort and depression/anxiety ratings (EQ5D5L), stress (PSS: 4-item perceived stress scale), resilience (BRCS: brief resilient coping scale), and somatic symptom disorder scoring (SSD-12: somatic symptom disorder - B criteria scale) were compared between participants with or without lung CT abnormality (A) and with or without diastolic dysfunction (B) at the one-year follow-up. Statistical significance was assessed by Mann-Whitney U test corrected for multiple testing with Benjamini-Hochberg method. Normalized (Z score) mean variance values with 2$\times$SEM are presented in radial plots. Numbers of complete observations are indicated under the plots._

\newpage

```{r fig-clust-qc, fig.width = mm_inch(suppl_figures$clust_qc$w), fig.height = mm_inch(suppl_figures$clust_qc$h), fig.cap = 'Development of COVID-19 recovery clusters.'}

suppl_figures$clust_qc$plot

```

__Supplementary Figure S\@ref(fig:fig-clust-qc). Development of COVID-19 recovery clusters.__ 

_Clustering of the study participants in respect to symptoms (any symptom present, dyspnea: modified Medical Research Council score [mMRC] $\geq$ 1, reduced performance: Eastern Cooperative Oncology Group score [ECOG] $\geq$ 1, fatigue: bimodal Chalder fatigue score [CFS] $\geq$ 4, self-reported: sleep problems, cough, night sweating and hyposmia/anosmia), mobility (6MWD < ref.: six minute walking distance, difference versus the reference value), cardiopulmonary abnormalities (any chest computed tomography [CT] abnormality: CT severity score $\geq$ 1, any lung function testing [LFT] abnormality, diastolic dysfunction), significant stress (4-item perceived stress scale [PSS] >5), impaired self-perceived general health (European quality of life 5 dimensions, 5 levels [EQ5D5L] visual analogue scale [VAS] <73, imp.: impaired) as well as features of quality of life and mental health (EQ5D5L dimensions, cutoff: score >1) at the one-year follow-up. Clustering analysis was done with the PAM algorithm (PAM: partitioning around medoids, simple matching distance [SMD])._

_(A) Performance comparison of various clustering algorithms and distance measures (algorithms: PAM, HCL: hierarchical clustering, SOM+PAM: combined self-organizing map/PAM, SOM+HCL: combined SOM/HCL algorithm, distances: Tanimoto and SMD). Clustering algorithm stability was assessed by the rate of correct cluster assignment in 10-fold cross-validation (CV). Clustering variance is expressed as a ratio of the total between-cluster sum of squares to total sum of squares. Based on the optimal stability, PAM (partitioning around medoids) algorithm with the SMD metric was chosen for further analyses._

_(B) The total within-cluster sum sum of squares as a function of the cluster number for the PAM/SMD clustering procedure. The optimal cluster number (dashed red line) was set at the bend of the curve._

_(C) Permutation importance of the clustering factors was determined by calculating differences in clustering variances of the original clustering structure with clustering objects generated with the particular clustering factors re-shuffled randomly ($\Delta$ clustering variance). Results for 100 random permutations (runs) are presented. Each point represents a separate run. Box plots represent median differences in clustering variance with interquartile ranges (IQR). whiskers span over 150% IQR._

\newpage

```{r fig-clust-relapse, fig.width = mm_inch(suppl_figures$clust_relapse$w), fig.height = mm_inch(suppl_figures$clust_relapse$h), fig.cap = 'Heat map of clustering features, frequency of CT abnormalities and rates of symptom relapse at the 12-month follow-up in the COVID-19 recovery clusters.'}

suppl_figures$clust_relapse$plot

```

__Supplementary Figure S\@ref(fig:fig-clust-relapse). Heat map of clustering features, frequency of CT abnormalities and rates of symptom relapse at the 12-month follow-up in the COVID-19 recovery clusters.__ 

_(A) Cardiopulmonary (CP) abnormalities (CT abnormality: CT severity score $\geq$ 1, any LFT abnormality, diastolic dysfunction), mobility impairment (6MWD > ref.: six minute walking distance, difference versus the reference value), symptoms (any symptom present, dyspnea: modified Medical Research Council score [mMRC] $\geq$ 1, reduced performance: Eastern Cooperative Oncology Group score [ECOG] $\geq$ 1, fatigue: bimodal Chalder fatigue score [CFS] $\geq$ 4, self-reported: sleep problems, cough, night sweating and hyposmia/anosmia), significant stress (4-item perceived stress scale [PSS] >5), impaired self-perceived general health (European quality of life 5 dimensions, 5 levels [EQ5D5L] visual analogue scale [VAS] <73, imp.: impaired) as well as features of quality of life and mental health (EQ5D5L dimensions, cutoff: score >1) at the one-year follow-up in the COVID-19 recovery clusters. Presence/absence of the clustering features is presented in a heat map. Numbers of participants assigned to the clusters are displayed next to the plot._

_(B) Frequencies of any chest CT abnormality at the one-year follow-up in ambulatory, moderate and severe COVID-19 patients assigned to the COVID-19 recovery clusters. Percentages of complete observations in each cluster are presented in bar plots. Numbers of complete observations are indicated in the Y axis._

_(C) Rates of relapse of COVID-19-related symptoms at the one-year follow-up as compared with the 6-month follow-up in the COVID-19 recovery clusters were assessed by $\chi^2$ test with Cramer V effect size statistic. Only participants with the complete symptom record for the investigated follow-up were included in the analysis. P values were corrected for multiple testing with Benjamini-Hochberg method. Percentages of the relapsing individuals are presented as a bar plot. Effect size statistics and p values for the significant comparisons are indicated in the plot. Numbers of participants assigned to the clusters are presented under the plot._

\newpage

```{r fig-clust-extra, fig.width = mm_inch(suppl_figures$clust_extra$w), fig.height = mm_inch(suppl_figures$clust_extra$h), fig.cap = 'Demographic features, COVID-19 severity, physical performance and mobility in the COVID-19 recovery clusters.'}

suppl_figures$clust_extra$plot

```

__Supplementary Figure S\@ref(fig:fig-clust-extra). Demographic features, COVID-19 severity, physical performance and mobility in the COVID-19 recovery clusters.__ 

_Age at COVID-19 diagnosis, sex, weight class distribution at COVID-19 diagnosis, acute COVID-19 severity, six-minute walking distance (6MWD vs ref.: difference between the observed and reference value) and physical performance scoring (ECOG: Eastern Cooperative Oncology Group score) at the one-year follow-up were compared between the COVID-19 recovery clusters. Statistical significance for numeric values was assessed by Kruskal-Wallis test with $\eta^2$ effect size statistic or by $\chi^2$ test with Cramer V effect size statistic for categorical variables. P values were corrected for multiple testing with Benjamini-Hochberg method. Effect size statistic and p values are presented in plot captions. Numbers of participants assigned to the clusters are presented under the plots._

\newpage

```{r fig-sympt-risk, fig.width = mm_inch(suppl_figures$sympt_risk$w), fig.height = mm_inch(suppl_figures$sympt_risk$h), fig.cap = 'Modeling of the persistent symptom risk at the one-year post-COVID-19 follow-up.'}

suppl_figures$sympt_risk$plot

```

__Supplementary Figure S\@ref(fig:fig-sympt-risk). Modeling of the persistent symptom risk at the one-year post-COVID-19 follow-up.__ 

_The risk of presence of COVID-19-related symptoms at the one-year follow-up was modeled with multi-parameter logistic LASSO (least absolute shrinkage and selection operator) regression and `r length(multi_mod$variables)` independent variables recorded during acute COVID-19 and at the two-month follow-up (FUP). Study participants with the complete independent variable set were included in the analysis, numbers of complete observations and participants with symptoms at the one-year follow-up are presented in A._

_(A) Non-zero model coefficient values presented as odds ratios (OR). Point size codes for the absolute OR value, point color codes for the correlation with the risk (blue: favorable, red: unfavorable factor). Time point of variable recording and number of participants in the variable strata are indicated in the Y axis._

_(B) Performance of the LASSO model at predicting persistent symptoms at the one-year follow-up in the training data set and 10-fold cross-validation (CV) assessed by receiver-operating characteristic (ROC). Area under the ROC curve (AUC), sensitivity (Se) and specificity (Sp) values with 95$\%$ confidence intervals (for CV) are presented in the plots. R-squared and Cohen's kappa statistic values are shown in the plot captions._

_NT-proBNP: N-terminal pro-brain natriuretic peptide; GID: gastrointestinal disease; #: number of; HM: hospitalized moderate COVID-19; IL6: interleukin 6._

\newpage

```{r fig-lft-risk, fig.width = mm_inch(suppl_figures$lft_risk$w), fig.height = mm_inch(suppl_figures$lft_risk$h), fig.cap = 'Modeling of the persistent functional lung abnormality at the one-year post-COVID-19 follow-up.'}

suppl_figures$lft_risk$plot

```

__Supplementary Figure S\@ref(fig:fig-lft-risk). Modeling of the persistent functional lung abnormality at the one-year post-COVID-19 follow-up.__ 

_The risk of any lung function testing (LFT) abnormality at the one-year follow-up was modeled with multi-parameter logistic LASSO (least absolute shrinkage and selection operator) regression and `r length(multi_mod$variables)` independent variables recorded during acute COVID-19 and at the two-month follow-up (FUP). Study participants with the complete independent variable set were included in the analysis, numbers of complete observations and participants with LFT abnormality at the one-year follow-up are presented in A._

_(A) Non-zero model coefficient values presented as odds ratios (OR). Point size codes for the absolute OR value, point color codes for the correlation with the risk (blue: favorable, red: unfavorable factor). Time point of variable recording and number of participants in the variable strata are indicated in the Y axis._

_(B) Performance of the LASSO model at predicting LFT abnormality at the one-year follow-up in the training data set and 10-fold cross-validation (CV) assessed by receiver-operating characteristic (ROC). Area under the ROC curve (AUC), sensitivity (Se) and specificity (Sp) values with 95$\%$ confidence intervals (for CV) are presented in the plots. R-squared and Cohen's kappa statistic values are shown in the plot captions._

_IL6: interleukin 6; #: number of._

\newpage

```{r fig-ct-risk, fig.width = mm_inch(suppl_figures$ct_risk$w), fig.height = mm_inch(suppl_figures$ct_risk$h), fig.cap = 'Modeling of the persistent radiological lung abnormality at the one-year post-COVID-19 follow-up.'}

suppl_figures$ct_risk$plot

```

__Supplementary Figure S\@ref(fig:fig-ct-risk). Modeling of the persistent radiological lung abnormality at the one-year post-COVID-19 follow-up.__ 

_The risk of any chest computed tomography (CT) abnormality (CT severity score $\geq$ 1) at the one-year follow-up was modeled with multi-parameter logistic LASSO (least absolute shrinkage and selection operator) regression and `r length(multi_mod$variables)` independent variables recorded during acute COVID-19 and at the two-month follow-up (FUP). Study participants with the complete independent variable set were included in the analysis, numbers of complete observations and participants with CT abnormality at the one-year follow-up are presented in A._

_(A) Non-zero model coefficient values presented as odds ratios (OR). Point size codes for the absolute OR value, point color codes for the correlation with the risk (blue: favorable, red: unfavorable factor). Time point of variable recording and number of participants in the variable strata are indicated in the Y axis._

_(B) Performance of the LASSO model at predicting CT abnormality at the one-year follow-up in the training data set and 10-fold cross-validation (CV) assessed by receiver-operating characteristic (ROC). Area under the ROC curve (AUC), sensitivity (Se) and specificity (Sp) values with 95$\%$ confidence intervals (for CV) are presented in the plots. R-squared and Cohen's kappa statistic values are shown in the plot captions._

_anti-S1/S2 IgG: anti-S1/S2 SARS-CoV-2 immunoglobulin; Q3, Q4: 3^rd^ and 4^th^ quartile; CRP: C-reactive protein; IL6: interleukin 6; HS: hospitalized severe COVID-19; #: number of._

\newpage

```{r fig-dysf-risk, fig.width = mm_inch(suppl_figures$dysf_risk$w), fig.height = mm_inch(suppl_figures$dysf_risk$h), fig.cap = 'Modeling of the persistent diastolic dysfunction at the one-year post-COVID-19 follow-up.'}

suppl_figures$dysf_risk$plot

```

__Supplementary Figure S\@ref(fig:fig-dysf-risk). Modeling of the persistent diastolic dysfunction at the one-year post-COVID-19 follow-up.__ 

_The risk of diastolic dysfunction at the one-year follow-up was modeled with multi-parameter logistic LASSO (least absolute shrinkage and selection operator) regression and `r length(multi_mod$variables)` independent variables recorded during acute COVID-19 and at the two-month follow-up (FUP). Study participants with the complete independent variable set were included in the analysis, numbers of complete observations and participants with diastolic dysfunction at the one-year follow-up are presented in A._

_(A) Non-zero model coefficient values presented as odds ratios (OR). Point size codes for the absolute OR value, point color codes for the correlation with the risk (blue: favorable, red: unfavorable factor). Time point of variable recording and number of participants in the variable strata are indicated in the Y axis._

_(B) Performance of the LASSO model at predicting diastolic dysfunction at the one-year follow-up in the training data set and 10-fold cross-validation (CV) assessed by receiver-operating characteristic (ROC). Area under the ROC curve (AUC), sensitivity (Se) and specificity (Sp) values with 95$\%$ confidence intervals (for CV) are presented in the plots. R-squared and Cohen's kappa statistic values are shown in the plot captions._

_GID: gastrointestinal disease; HS: hospitalized severe COVID-19; CRP: C-reactive protein; HM: hospitalized moderate COVID-19; FT: ferritin; CVD: cardiovascular disease; anti-S1/S2 IgG: anti-S1/S2 SARS-CoV-2 immunoglobulin; Q3 and Q4: 3^rd^ and 4^th^ quartile._

\newpage

# References
